### 2020-01-31 paper/sex_selection_spacing-ver3.tex
No longer a need for a long descriptive stats description

The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit, the birth of a child.%
\footnote{
See \citet{Sheps1970} and \citet{Newman1984} for early discussions of why 
hazard models are the preferred way to deal with the censoring of birth
intervals.
}
There are, however, two problems with the standard approach in this setting.
First, and most importantly, the introduction of sex selection means that the sex of the 
next child is no longer necessarily random and that parents' choices will impact the 
spacing to the birth of a girl or a boy differently.
I, therefore, use a competing risk set-up, which can capture both the non-randomness of
the birth outcome and the differential spacing.
To my knowledge, this is the first application of competing risk models to birth
spacing.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

@@

\subsection{Descriptive Statistics}

@@

The share of urban women in the sample has fallen slightly over the
periods, even though India's population has become more urbanized, presumably because the 
age of marriage has increased faster in urban than in rural areas.
Hence, there are relatively more urban women, but fewer show up in the
sample because they are not yet married.

Finally, the population has become substantially better educated over time.
In the first period, almost 60\% of women had no education, and only twenty percent had
eight or more years, while in the last period the reverse was the case.
These changes underestimate the increase in education because many of the younger women 
with more education are not yet married and therefore not in the sample.%
\footnote{
Women with eight or more years of education accounted for 65.9\% of
unmarried women in NFHS-3 and 82.8\% of unmarried women in NFHS-4
\citep{International-Institute-for-Population-Sciences-IIPS2007,International-Institute-for-Population-Sciences-IIPS2017}.
}

% \citep[p 56]{International-Institute-for-Population-Sciences-IIPS2007}
% and \citet[p 61]{International-Institute-for-Population-Sciences-IIPS2017}
% for more information.

% Age at marriage increased over time across all three education groups.
% The most substantial increase was for women without any education, where
% the average went from below 16 years of age to 18.5.
% The smallest change is for the most educated women where the average age
% at marriage went up by less than a year---from 19.6 to 20.4---across the 
% four periods.

@@

Instead of averaging across the entire sample, for each spell/education combination I 
calculate the conditional survival curves using average values and the majority of
categories, which means no ownership of land and not in a scheduled caste or tribe.
The characteristics used do not change across the four decades to ensure that composition 
effects do not drive the changes.
In the interest of brevity, I discuss only two subsamples.
Figures \ref{fig:pps_low} and \ref{fig:pps_high} show spacing across sex compositions for 
the second, third, and fourth spells for the first period, 1972--1984, and the last 
period, 2005--2016, for rural women with no education and urban women with eight or 
more years of education.
The Appendix shows conditional survival curves for all groups and decades and to 
complement this approach, Appendix Tables \ref{tab:p25_p50_p75_low}, 
\ref{tab:p25_p50_p75_med}, and \ref{tab:p25_p50_p75_high} show 25th, 50th, and 75th 
percentiles durations with bootstrapped standard errors based on averages across all women.

@@

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\footnotesize{Second Spell}}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{0}
\subfloat[2005--2016]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\footnotesize{Third Spell}}
\setcounter{subfigure}{1}
\subfloat[1972--1984]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
% \\
% \rotatebox[origin=c]{90}{\footnotesize{Fourth Spell}}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.46\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.46\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{
% Survival curves conditional on progression to next birth for urban women with eight or
% more years of education; start point is nine months after prior birth
% }
\label{fig:pps_high}
\end{figure}

>>>>

The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit, the birth of a child.%
\footnote{
See \citet{Sheps1970} and \citet{Newman1984} for early discussions of why 
hazard models are the preferred way to deal with the censoring of birth
intervals.
}
There are, however, two problems with the standard approach in this setting.
First, and most importantly, the introduction of sex selection means that the sex of the 
next child is no longer necessarily random and that parents' choices will impact the 
spacing to the birth of a girl or a boy differently.
I, therefore, use a competing risk set-up, which can capture both the non-randomness of
the birth outcome and the differential spacing.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

@@

@@

@@

Instead of averaging across the entire sample, for each spell/education combination I 
calculate the conditional survival curves using average values and the majority of
categories, which means no ownership of land and not in a scheduled caste or tribe.
The characteristics used do not change across the four decades to ensure that composition 
effects do not drive the changes.
In the interest of brevity, I discuss only two subsamples.
Figures \ref{fig:pps_low} and \ref{fig:pps_highest} show spacing across sex compositions for 
the second, third, and fourth spells for the first period, 1972--1984, and the last 
period, 2005--2016, for rural women with no education and urban women with eight or 
more years of education.
The Appendix shows conditional survival curves for all groups and decades and to 
complement this approach, Appendix Tables \ref{tab:p25_p50_p75_low}, 
\ref{tab:p25_p50_p75_med}, and \ref{tab:p25_p50_p75_high} show 25th, 50th, and 75th 
percentiles durations with bootstrapped standard errors based on averages across all women.

@@

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\footnotesize{Second Spell}}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{0}
\subfloat[2005--2016]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\footnotesize{Third Spell}}
\setcounter{subfigure}{1}
\subfloat[1972--1984]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.46\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
% \\
% \rotatebox[origin=c]{90}{\footnotesize{Fourth Spell}}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.46\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.46\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Probability of no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
\caption{
Survival curves conditional on progression to next birth for urban women with twelve or
more years of education; start point is nine months after prior birth
}
\label{fig:pps_highest}
\end{figure}
